\chapter{Introduction}
\label{ch:introduction}

\section{Context and Motivation}

\textbf{Template intent.} Explain why Formula 1 pit-wall decision support is a strong case for streaming analytics plus machine learning.

\begin{itemize}
	\item Start from the operational problem: decisions are sequential, time-critical, and made under uncertainty.
	\item Explain why batch-only analysis is insufficient for race-time decisions.
	\item Position event-time processing and low-latency architecture as enabling conditions.
\end{itemize}

Suggested references: \citep{akidau2015dataflow, carbone2015flink, kreps2011kafka, widanage2019indy500}.

\section{Research Scope and Objectives}

\textbf{Template intent.} Define exactly what the thesis will and will not do.

\begin{itemize}
	\item Scope: pit-stop recommendation as a supervised lap-level decision task in a streaming architecture.
	\item Comparison target: deterministic heuristic strategy engine vs learned model on the same replay stream.
	\item Explicit non-goals: full race-winning simulation, proprietary telemetry replication, and team-specific private strategy rules.
	\item Add 2--4 research questions, each measurable in Chapter 6.
\end{itemize}

Suggested references: \citep{sasikumar2025pitstop, hettmann2024datatopodium, dinh2025gradient, thomas2025xrlf1}.

\section{Contributions}

\textbf{Template intent.} List the contribution claims in a thesis-safe and testable format.

\begin{itemize}
	\item A reproducible event-time data engineering pipeline for F1 replay analytics.
	\item A leakage-aware ML protocol for pit-stop recommendation.
	\item A fair comparison protocol between heuristic and ML decisions in identical operating conditions.
	\item A reproducibility package (data contracts, scripts, environment, and experiment documentation).
\end{itemize}

Suggested references: \citep{carrasco2023optimization, quiroga2024optimization, piccinotti2020openloop}.

\section{Thesis Organization}

\textbf{Template intent.} Provide a concise chapter roadmap.

\begin{itemize}
	\item Chapter 2: state of the art and research gap.
	\item Chapter 3: problem definition, data, labels, and evaluation metrics.
	\item Chapter 4: solution design and comparison protocol.
	\item Chapter 5: implementation choices and reproducibility.
	\item Chapter 6: evaluation campaign and discussion.
	\item Chapter 7: conclusions, limitations, and future work.
\end{itemize}
