\chapter{Solution and Design}
\label{ch:solution_and_design}

\section{System Architecture Overview}

\textbf{Template intent.} Describe end-to-end architecture in one page.

\begin{itemize}
	\item Producer, broker, stream processor, data lake, and dashboard roles.
	\item Explain how architecture choices support reproducibility and fair comparisons.
\end{itemize}

Suggested references: \citep{akidau2015dataflow, carbone2015flink, kreps2011kafka}.

\section{Dataflow and Processing Topology}

\textbf{Template intent.} Specify operator order and data dependencies.

\begin{itemize}
	\item Event-time path from ingestion to features/alerts/labels.
	\item Stateful operators and where context enters the pipeline.
	\item Contracts for sinks used by audit and ML.
\end{itemize}

\subsection{Dual-Consumer Streaming Contract (Flink + ML Placeholder)}

\textbf{Template intent.} Reserve this subsection for deployment-consistency design notes.

\begin{itemize}
	\item State contract: Flink computes stateful lap-context features once and emits the same feature payload to offline sink and online topic (\texttt{f1-ml-features}).
	\item Document event-time ordering and watermark assumptions used to keep feature timestamps point-in-time consistent across replay and live serving.
	\item Describe Kafka decoupling between heuristic processing and ML inference consumer to preserve independent runtime behavior under the same telemetry stream.
	\item Keep raw-telemetry serving as a deferred ablation track after consistency and fairness gates are validated.
\end{itemize}

Suggested references: \citep{carbone2015flink, akidau2015dataflow, kreps2011kafka, widanage2019indy500}.

\section{Heuristic Strategy Engine Design}

\textbf{Template intent.} Explain the handcrafted strategy logic at a conceptual level.

\begin{itemize}
	\item Inputs, decision factors, and prioritization logic.
	\item Expected strengths and known failure modes.
	\item Clarify this module as baseline and operational comparator.
\end{itemize}

\subsection{Pit-Cycle Label Resolution (Skeleton Notes)}

\textbf{Template intent.} Keep this subsection structural and implementation-neutral in early drafts.

\begin{itemize}
	\item Document fallback hierarchy order: direct rival-gap path, near-lap recovery, pace-shift fallback, unresolved abstention.
	\item Explain why conservative unresolved outcomes are preserved when evidence remains incomplete.
	\item Link rationale to lap-time decomposition and online optimization framing.
\end{itemize}

Suggested references: \citep{heilmeier2020race, carrasco2023optimization, quiroga2024optimization}.

\subsection{Timing-Pressure Actionability Alignment (Skeleton Notes)}

\textbf{Template intent.} Record purpose and guardrails of timing-aware suggestion logic without locking final numeric detail.

\begin{itemize}
	\item Summarize timing-pressure signals (pace trend, gap trend, strategy deficit, rival reaction, traffic context).
	\item Note green-condition actionability gating that downgrades weak early actionable alerts to monitoring.
	\item Capture near-window re-emission intent when timing pressure materially increases.
	\item Emphasize deterministic bounded behavior to keep the SDE baseline stable and auditable.
\end{itemize}

Suggested references: \citep{widanage2019indy500, heilmeier2020race, carrasco2023optimization, quiroga2024optimization, cappello2025statespace}.

\section{Machine Learning Formulation}

\textbf{Template intent.} Specify model-family choices and data protocol without reporting outcomes yet.

\begin{itemize}
	\item Candidate baselines (for example logistic/XGBoost and optional sequence model).
	\item Feature groups, leakage controls, and temporal/grouped split strategy.
	\item Traffic-context family for pit feasibility, integrated from drop-zone state (positions lost and nearest physical-car gap) on race-driver-lap keys.
	\item Causal momentum family based on lagged differences (for example $k-2$ pace and gap trends) to preserve point-in-time availability.
	\item Calibration and thresholding plan for decision outputs.
\end{itemize}

Suggested references: \citep{hettmann2024datatopodium, dinh2025gradient, sasikumar2025pitstop, carrasco2023optimization, quiroga2024optimization, widanage2019indy500, heilmeier2020race, brookshire2024leakage, roberts2017cvstructured, platt1999probabilistic, guo2017calibration, kull2017beta}.

\section{Comparison Protocol, Heuristic vs ML}

\textbf{Template intent.} Define a fairness protocol before experiments.

\begin{itemize}
	\item Same replay windows, same data availability assumptions, same excluded contexts.
	\item Fairness contract is frozen in this cycle: actionable-only matching, one-to-one pit consumption, fixed horizon $H=2$.
	\item Raw telemetry branch is deferred to Phase 3.2, comparator semantics are not relaxed in Phase 3.1 claims.
	\item Use one shared feature contract for training and first live serving comparison to reduce training-serving skew.
	\item Treat decision threshold as an operating-point control variable, selected from a precision-coverage frontier under a predefined safety floor.
	\item Report both predictive quality and operational behavior.
	\item Predefine significance testing and uncertainty reporting.
\end{itemize}

Suggested references: \citep{dietterich1998approximate, nadeau2003inference, demsar2006statistical, walters2022comparing, elkan2001costsensitive}.
