\chapter{Conclusion}
\label{ch:conclusion}

\section{Obtained Results}

\textbf{Template intent.} Keep this section high-level until final numbers are fixed.

\begin{itemize}
	\item Delivered an end-to-end comparison pipeline under a fixed fairness contract ($H=2$, actionable-only, one-to-one matching) between deterministic SDE and engineered-feature ML.
	\item Merged 2022--2024 evidence shows a precision gain at constrained ML operating policy: SDE precision $0.7213$ vs ML precision $0.9413$, with independent scored-row test reported as $p=0.0$ (numerical underflow).
	\item Threshold frontier analysis confirms controllable precision-coverage tradeoff; threshold $0.10$ remains the balanced reporting point (precision $0.8726$, above the SDE floor $0.7213$).
	\item Integrated readiness gate returned GO (high confidence): all core phases B/C/D/F/G passed, with $12/12$ by-layer checks passing, parity $1.0000$, latency $p95=7.758$ ms, and availability $100.0\%$.
	\item Session-level hardening completed before this freeze: duplicate-lap keep-last serving semantics, pre-transform dedup in parity replay, merged \texttt{ml\_features} dedup validation in evaluator, and dynamic integrated gate note attribution.
\end{itemize}

\section{Limitations}

\textbf{Template intent.} Be explicit and concrete.

\begin{itemize}
	\item Comparator scope limitation: fixed $H=2$ fairness contract can under-credit decisions whose value appears beyond the horizon.
	\item Label-policy limitation: unresolved and weather-sensitive contexts are conservatively excluded, improving purity but reducing effective coverage.
	\item Statistical pairing limitation: McNemar overlap support is small, so paired significance interpretation is reported as sensitivity evidence.
	\item Deployment limitation: integrated Phase H gate returned GO for the engineered-feature scope, but claims remain bounded to replay-protocol evidence and fixed fairness-contract assumptions.
	\item Contract fragility limitation: parity depends on stable keep-last de-duplication order in merged feature streams; upstream key or ordering changes can re-introduce silent train-serve skew.
	\item Operational limitation: drift/skew monitoring and long-horizon live fallback stress tests remain future hardening work beyond this cycle.
\end{itemize}

\section{Future Work}

\textbf{Template intent.} Propose realistic next steps tied to thesis findings.

\begin{itemize}
	\item Drift-aware extension: add season-shift monitoring, recalibration policy triggers, and sustained fallback stress testing under live operations.
	\item Comparator robustness extension: institutionalize split-integrity and comparator-invariance audits as recurring gate checks in future merged reruns.
	\item Deferred Phase 3.2 branch: evaluate raw telemetry formulations only after explicit Phase I entry prerequisites are met.
	\item Systems extension: run Kafka-in-loop latency sensitivity checks to complement replay-based operational baselines.
\end{itemize}

Suggested references: \citep{gama2014drift, arora2024driftreview, bifet2007adwin, gomes2017arf, baenagarcia2006eddm}.

\subsection{Deployment Consistency Placeholder (Flink + ML)}

\textbf{Template intent.} Keep this subsection as a future-work marker until deployment experiments are run.

\begin{itemize}
	\item Keep one-feature-authority policy: Flink emits serving and offline feature payloads from the same event-time logic.
	\item Keep parity checklist explicit: schema parity, point-in-time validity, sampled offline-vs-live consistency.
	\item Keep operational checklist explicit: prediction availability, fallback behavior, and latency budgets, now closed for this cycle through Phase H GO evidence.
\end{itemize}

\subsection{Phase I Scope Lock for Deferred Raw Telemetry Branch}

\textbf{Template intent.} Keep raw telemetry work explicitly deferred and bounded with measurable entry criteria.

\begin{itemize}
	\item Scope lock for this cycle: engineered-feature ML only, with raw telemetry modeling deferred to Phase 3.2.
	\item Entry prerequisites for Phase 3.2 remain explicit and auditable: integrated gate state not weaker than GO WITH CONDITIONS, stable drift/skew monitoring protocol, and documented point-in-time raw feature extraction with leakage audits.
	\item Method guardrail for future Phase 3.2 experiments: preserve grouped temporal validation and fairness-contract transparency while treating construct-validity caveats explicitly.
	\item Phase I lock artifacts: \texttt{phase\_i\_scope\_lock\_2022\_2024\_merged.txt}, \texttt{phase\_i\_future\_work\_prerequisites\_2022\_2024\_merged.md}.
\end{itemize}

Suggested references: \citep{brookshire2024leakage, roberts2017cvstructured, gama2014drift, arora2024driftreview, widanage2019indy500, cappello2025statespace, heilmeier2020race}.

\section{Overall Discussion}

\textbf{Template intent.} Close by positioning your thesis against the three strategy paradigms.

\begin{itemize}
	\item Optimization/planning paradigm remains a strategic reference for global race planning and constraint reasoning.
	\item Reinforcement-learning paradigm remains a deferred comparator pending broader online evaluation infrastructure.
	\item Supervised classification in streaming deployment is the validated contribution of this cycle, with significance-backed precision gains under explicit fairness and reproducibility constraints.
\end{itemize}

Suggested references: \citep{carrasco2023optimization, quiroga2024optimization, thomas2025xrlf1, sasikumar2025pitstop, hettmann2024datatopodium, dinh2025gradient}.
